The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery

### 2021-10-20 adaptation/article/article.tex
Small changes on article

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}. Here, we use this principle and chose to make use of an exact Bayesian learner. As has been discussed before \cite{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminating the need for explicit multiple time scales of learning as the learner can deal with adaptation after any number of trials. As we will show below this feature will be useful to explain specific key phenomena parsimoniously.

@@

\subsection{Savings and slow/fast switching}
The term 'savings' refer to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it. Savings is almost universally observed in human participants \cite{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}. In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.

@@

Because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. In fact, the mere act of holding the handle informs participants that something is different from baseline. Proprioceptive signals can also provide contextual information \cite{Dizio_Motor_1995,Shadmehr_Adaptive_1994}. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite{Ethier_Spontaneous_2008}.

To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies \cite{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information. \cite{Kim_Neural_2015} performed a cued-context visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees. Participants performed shooting movements in blocks of trials with the same rotation. Importantly, the identity of the current context was displayed as a visual cue (colored light), making this a cued-context experiment. Consistent with our model, the authors found that switching was immediate and accurate. In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}. The correspondence between simulations and experimental results can be seen in the switches between contexts, i.e. when the solid black line (representing the true context) switches between 40, 0 and -40; in both panels (left and right), the blue line, representing the agent's adaptation, quickly follows the switches.

>>>>

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}. Here, we use this principle and chose to make use of an exact Bayesian learner. As has been discussed before \citep{Oh_Minimizing_2019}, a Bayesian learner has the advantage of not only fitting experimental data on motor adaptation, but also eliminating the need for explicit multiple time scales of learning as the learner can deal with adaptation after any number of trials. As we will show below this feature will be useful to explain specific key phenomena parsimoniously.

@@

\subsection{Savings and slow/fast switching}
The term 'savings' refer to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it. Savings is almost universally observed in human participants \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}. In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.

@@

Because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. In fact, the mere act of holding the handle informs participants that something is different from baseline. Proprioceptive signals can also provide contextual information \citep{Dizio_Motor_1995,Shadmehr_Adaptive_1994}. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \citep{Ethier_Spontaneous_2008}.

To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies \citep{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information. \cite{Kim_Neural_2015} performed a cued-context visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees. Participants performed shooting movements in blocks of trials with the same rotation. Importantly, the identity of the current context was displayed as a visual cue (colored light), making this a cued-context experiment. Consistent with our model, the authors found that switching was immediate and accurate. In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results by \cite{Kim_Neural_2015}. The correspondence between simulations and experimental results can be seen in the switches between contexts, i.e. when the solid black line (representing the true context) switches between 40, 0 and -40; in both panels (left and right), the blue line, representing the agent's adaptation, quickly follows the switches.
